---
title: "Reproducible Research: Peer Assessment 1"
output: 
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading and preprocessing the data

## load the packages
```{r message=FALSE}
library (data.table)
library(dplyr)
library(ggplot2)
directory <- getwd()
```

## load the data 
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = paste0(getwd(), '/activity.zip'), method = "curl")
unzip("activity.zip",exdir = "data")
activitydata <- read.csv(paste0(directory,'/data/activity.csv'))
```


# What is mean total number of steps taken per day?

## group data by date
```{r}
stepday <- activitydata %>% group_by(date) %>% summarize(sum_steps = sum(steps), .groups = 'drop' )
```

## make a histogram of the total numbers of steps 
```{r}
with(
  stepday,
  hist(
    sum_steps,
    xlab = "no. of steps per day",
    labels = TRUE
  )
)
```
  
## calculate mean and median 

```{r}
stepday %>% filter(complete.cases(.)) %>% summarize (mean= mean(sum_steps), median = median (sum_steps))
```


## What is the average daily activity pattern?
```{r}
activityinterval <- activitydata %>% group_by(interval) %>% na.omit () %>% 
  summarize (mean_step = mean(steps), .groups='drop')

ggplot(activityinterval, aes(interval, mean_step)) + geom_line() + 
  labs(x= "interval", y= "avg steps") +
  ggtitle("avg step during 5min interval")

```
  
## which 5min interval contains max steps? 

```{r}
activityinterval %>% filter(mean_step == max(mean_step))
```


# Imputing missing values
## calculate total number of NAs
```{r}
activitydata %>% summarise_all(~sum(is.na(.)))
```

## fill in all NA : use the mean of the day,  create a new dataset that has data filled in 
```{r}
completedata <- activitydata
for (i in 1:nrow(completedata)) {
  if(is.na(completedata$steps[i])) {
    index <- which(completedata$interval[i]== activityinterval$interval)
    completedata$steps[i] <- activityinterval[index, ]$mean_step
    
  }
}
```

## make histogram and calculate mean and median 

```{r}
plot_completedata <- completedata %>% 
  group_by(date) %>% 
  summarize(sum_steps = sum(steps), .groups= 'drop')

with(
  plot_completedata,
  hist(
    sum_steps,
    xlab= "no. of steps per day",
    labels= TRUE
  )
) 
```

# Are there differences in activity patterns between weekdays and weekends?

## divide into weekend and weekdays 
```{r}
head(completedata)
completedata$day <- as.factor(weekdays(as.Date(completedata$date)))
completedata$is_weekday <- ifelse(!(completedata$day %in% c("Saturday","Sunday")), TRUE, FALSE)
```

## mean of weekday 
```{r}
wkd_dt <- completedata[completedata$is_weekday,]
wkd_mean <- aggregate(wkd_dt$steps, by=list(interval=wkd_dt$interval), FUN=mean)
```

## mean of weekday 
```{r}
wkend_dt <-completedata[!completedata$is_weekday,]
wkend_mean <- aggregate(wkend_dt$steps, by=list(interval=wkend_dt$interval),FUN=mean)
```

## name  columns and plot
```{r}
colnames(wkd_mean)<-c("interval","avg_steps")
colnames(wkend_mean) <-c("interval","avg_steps")
wkd_mean$day <- "Weekday"
wkend_mean$day <- "Weekend"

all_week <- rbind(wkd_mean,wkend_mean)
all_week$day <- as.factor(all_week$day)
library(lattice)
xyplot(avg_steps ~ interval | day, data= all_week, layout = c(1,2), type="l",xlab = "interval",  ylab="no. of steps")
```